Structure Learning for $\VAR(p)$ Models

### 2019-05-29 paper/paper.tex
Some results and more code for simulations.

%% bibliography stuff -- this needs come before the preamble inclusion
\usepackage[backend=bibtex,sorting=none]{biblatex}
\usepackage{enumitem}
\usepackage{etex,etoolbox}
\bibliography{\string~/Documents/academics/global_academics/global_bib.bib}
\usepackage{hyperref}

@@

\begin{proposition}
  Algorithm \ref{alg:pwgr_heuristic} terminates in at most $n$ iterations and returns a strongly causal graph $\widehat{\gcg}$.
\end{proposition}
\begin{proof}
  \hl{TODO}
\end{proof}

A complete estimation procedure for a sparse filter $\widehat{\B}(z)$
used to estimate $x(t + 1)$ from $\H_{t - 1}$ is then given in Algorithm \ref{alg:Bz_estimate}.

\begin{proposition}
  The error sequence in Algorithm \ref{alg:Bz_estimate} is a zero mean
  WSS process $\E\epsilon^{(m)}(t) = 0$ with monotonically decreasing
  variance: $\E|\epsilon^{(m + 1)}(t)|^2 \le \E|\epsilon^{(m)}(t)|^2$.
\end{proposition}
\begin{proof}
  \hl{TODO -- this is probably not even true}
\end{proof}

@@

  \While{$S \ne \emptyset$}{
    $S \leftarrow S \setminus P_{k - 1}$ \texttt{\# remove processed nodes}\\
    $\mathcal{I}_k \leftarrow \big(\sum_{j \in S: (j, i) \in W_\delta} F_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# intra-}$S$ \texttt{incident strength}\\
    $P_k \leftarrow \mathsf{arg\_select\_min\_N}(\mathcal{I}_k, N=b)$  \texttt{\# select} $b$ \texttt{branching nodes}\\
    \;
    \texttt{\# add strongest} $P_{k - r} \rightarrow P_k$ \texttt{edges and maintain strong causality}\\
    \For{$r = 1, \ldots, k$} 
    {
      \For{$(i, j) \in \mathsf{sort}\Big(\{(i, j) \in P_{k - r} \times P_k\ |\ (i, j) \in W_\delta\} \mathsf{\ by\ descending\ } F_{ji}\Big)$}{
        \If{\textit{no }$\gcgpath{i}{j}$\textit{ path in }$E$}{
          $E \leftarrow E \cup \{(i, j)\}$
        }
      }
    }
    $k \leftarrow k + 1$
  }
  \Return{$([n], E)$}
\end{algorithm}

@@

\section{Application}
\label{sec:application}
\section{Conclusion}
\label{sec:conclusion}
Extensions to transfer entropy?.  Case of time varying statistics?.  Develop sparse estimators for SS models (\cite{barnett2015granger})?.  Apply MIO for the Whittle relaxation?

>>>>

%% bibliography stuff -- this needs come before the preamble inclusion
\usepackage[backend=bibtex,sorting=none]{biblatex}
\usepackage{enumitem}
\usepackage{etex,etoolbox}
\usepackage{hyperref}
\usepackage{todonotes}
\bibliography{\string~/Documents/academics/global_academics/global_bib.bib}

@@

% \begin{proposition}
%   Algorithm \ref{alg:pwgr_heuristic} terminates in at most $n$ iterations and returns a strongly causal graph $\widehat{\gcg}$.
% \end{proposition}
% \begin{proof}
%   \hl{TODO}
% \end{proof}

% A complete estimation procedure for a sparse filter $\widehat{\B}(z)$
% used to estimate $x(t + 1)$ from $\H_{t - 1}$ is then given in Algorithm \ref{alg:Bz_estimate}.

% \begin{proposition}
%   The error sequence in Algorithm \ref{alg:Bz_estimate} is a zero mean
%   WSS process $\E\epsilon^{(m)}(t) = 0$ with monotonically decreasing
%   variance: $\E|\epsilon^{(m + 1)}(t)|^2 \le \E|\epsilon^{(m)}(t)|^2$.
% \end{proposition}
% \begin{proof}
%   \hl{TODO -- this is probably not even true}
% \end{proof}

@@

  \While{$S \ne \emptyset$}{
    $S \leftarrow S \setminus P_{k - 1}$ \texttt{\# remove processed nodes}\\
    $\mathcal{I}_k \leftarrow \big(\sum_{j \in S: (j, i) \in W_\delta} F_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# intra-}$S$ \texttt{incident strength}\\
    $P_k \leftarrow \mathsf{arg\_select\_min\_N}(\mathcal{I}_k, N=b)$  \texttt{\# select} $b$ \texttt{branching nodes}\\
    \;
    \texttt{\# add strongest} $P_{k - r} \rightarrow P_k$ \texttt{edges and maintain strong causality}\\
    \For{$r = 1, \ldots, k$} 
    {
      \For{$(i, j) \in \mathsf{sort}\Big(\{(i, j) \in P_{k - r} \times P_k\ |\ (i, j) \in W_\delta\} \mathsf{\ by\ descending\ } F_{ji}\Big)$}{
        \If{no $\gcgpath{u}{j}$ path in $E$, for any $u \in \anc{i}$}{
          $E \leftarrow E \cup \{(i, j)\}$
        }
      }
    }
    $k \leftarrow k + 1$
  }
  \Return{$([n], E)$}
\end{algorithm}

@@

\section{Empirical Evaluation}
We have implemented simple empirical experiments in Python
\cite{scipy}, in particular we leverage the LASSO implementation from
\texttt{sklearn} \cite{sklearn} and the random graph generators from
\texttt{networkx} \cite{networkx}.  We run the same experiment on
three separate graph topologies having $n = 75$ nodes: a tree (which
is strongly causal), a random directed acyclic graph, and an Erdos
Renyi random graph with edge inclusion probability of $2 / n$.

We populate each of the edges (including self loops) with random
linear filters constructed by placing $5$ transfer function poles
uniformly at random in a disc of radius $3 / 4$ (which guarantees
stability for acyclic graphs).  The resulting system is driven by
i.i.d. Gaussian random vectors, each component having random variance
$\sigma_i^2 = 1/2 + r_i$ where $r_i \sim \text{exp}(1/2)$.  Results
are obtained for $50$ random iterations per each of $6$ dataset
lengths $T \in \{100, 500, 1000, 2500, 5000, 10000\}$.

The pairwise causality function $\mathsf{pwgc}$ of Algorithm
\ref{alg:Bz_estimate} are obtained via the residuals
$\mathbf{\epsilon}_{ij}$ of the model
$\mathbf{x}_i = \mathbf{X}_{ij}\widehat{\beta_{}} +
\mathbf{\epsilon}_{ij}$ estimated via ordinary least squares, where
$\mathbf{X}_{ij}$ includes $10$ lagged values of processes $i$ and
$j$.

The function $\mathsf{estimate_B}$ in Algorithm \ref{alg:Bz_estimate}
is carried out via the LASSO \cite{tibshirani2015statistical}, with
the regularization parameter being chosen to minimize the Bayesian
information criteria, a more conservative heuristic than the AIC.

\begin{figure}
  \centering
  \caption{Simulation Results and Representitive Random Graph Topologies}
  \label{fig:simulation_results}

  \begin{subfigure}[b]{0.30\textwidth}
    \caption{Random Tree}
    \includegraphics[width=\linewidth]{example_tree.pdf}
  \end{subfigure}
  \begin{subfigure}[b]{0.30\textwidth}
    \caption{Random DAG}
    \includegraphics[width=\linewidth]{example_dag.pdf}
  \end{subfigure}
  \begin{subfigure}[b]{0.30\textwidth}
    \caption{Erdos-Renyi Graph}
    \includegraphics[width=\linewidth]{example_gnp.pdf}
  \end{subfigure}

  \begin{subfigure}[b]{0.30\textwidth}
    \includegraphics[width=\linewidth]{random_tree_simulation.pdf}
  \end{subfigure}
  \begin{subfigure}[b]{0.30\textwidth}
    \includegraphics[width=\linewidth]{random_dag_simulation.pdf}
  \end{subfigure}
  \begin{subfigure}[b]{0.30\textwidth}
    % \includegraphics[width=\linewidth]{random_gnp_simulation.pdf}
    \missingfigure{}
  \end{subfigure}
\end{figure}

\section{Application}
\label{sec:application}
\section{Conclusion}
\label{sec:conclusion}
% Extensions to transfer entropy?.  Case of time varying statistics?.  Develop sparse estimators for SS models (\cite{barnett2015granger})?.  Apply MIO for the Whittle relaxation?
